---
title: "p8105_hw6_yg3096"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
library(purrr)
library(modelr)    
library(p8105.datasets) 
data("weather_df")
set.seed(123)
```

## Problem 1
```{r}
# Read data
homi_raw <- readr::read_csv("data/homicide-data.csv", show_col_types = FALSE)

# Clean and process data
homi <- homi_raw %>%
        janitor::clean_names() %>%
        mutate(solved = if_else(disposition == "Closed by arrest", 1L, 0L),
               # Change unknown age into NA
               victim_age = na_if(as.character(victim_age), "Unknown"),
               victim_age = suppressWarnings(as.numeric(victim_age)),
               # White/Black only
               victim_race = case_when(victim_race == "White" ~ "White",
                                       victim_race == "Black" ~ "Black",
                                       TRUE ~ NA_character_),
               victim_sex = factor(victim_sex, levels = c("Female", "Male")),
               victim_race = factor(victim_race, levels = c("White", 
                                                            "Black"))) %>%
        # Filter cities
        filter(!(city == "Dallas" & state == "TX"),
               !(city == "Phoenix" & state == "AZ"),
               !(city == "Kansas City" & state == "MO"),
               !(city == "Tulsa" & state == "AL")) %>%      
        # Filter race
        filter(!is.na(victim_race))

# Data for baltimore
bal_df <- homi %>%
          filter(city == "Baltimore", state == "MD") %>%
          filter(!is.na(victim_age), !is.na(victim_sex), 
                 !is.na(victim_race)) %>%
          droplevels()

# Fit model for baltimore
bal_fit <- glm(solved ~ victim_age + victim_sex + victim_race, data = bal_df, 
               family = binomial())

# Odds ratio
bal_or <- tidy(bal_fit, exponentiate = TRUE, conf.int = TRUE) %>%
          filter(term == "victim_sexMale") %>%
          transmute(city_state = "Baltimore, MD", term, OR = estimate, 
                    CI_low = conf.low, CI_high = conf.high, p = p.value)
bal_or

# Helper function for glm
glm_function <- function(df) {
                df <- df %>%
                filter(!is.na(victim_age), !is.na(victim_sex), 
                       !is.na(victim_race)) %>%
                droplevels()
                
                if (nrow(df) == 0 || n_distinct(df$victim_sex) < 2 ||
                   n_distinct(df$victim_race) < 2) {
                   return(NULL)
                }

                fit <- glm(solved ~ victim_age + victim_sex + victim_race, 
                           data = df, family = binomial())
                tidy(fit, exponentiate = TRUE, conf.int = TRUE) %>%
                filter(term == "victim_sexMale") %>%
                transmute(OR = estimate, CI_low = conf.low, CI_high = conf.high, 
                          p = p.value)
}

# Fit for all cities
city_or <- homi %>%
           group_by(city, state) %>%
           group_split() %>%
           map_dfr(function(df) {
           this_city <- df$city[1]
           this_state <- df$state[1]
           res <- glm_function(df)
           if (is.null(res) || nrow(res) == 0) {
             tibble(city = this_city, state = this_state,
             OR = NA_real_, CI_low = NA_real_, CI_high = NA_real_, 
             p = NA_real_)
           } else {
             res %>% mutate(city = this_city, state = this_state, .before = 1)
           }
           }) %>%
           filter(!is.na(OR)) %>%
           mutate(city_state = paste(city, state, sep = ", ")) %>%
           arrange(OR)
head(city_or)

# Plot of odds ratio
plot <- city_or %>%
        mutate(city_state = forcats::fct_reorder(city_state, OR)) %>%
        ggplot(aes(x = city_state, y = OR)) +
        geom_pointrange(aes(ymin = CI_low, ymax = CI_high)) +
        coord_flip() + 
        labs(title = "Adjusted OR of solving homicides: Male vs Female by city",
             x = NULL, y = "Adjusted OR (Male vs Female)") 
plot
```

Across cities, the majority of 95% CIs covers 1, and about 20 cities sit below 1  
(male-victim cases less likely to be solved). This heterogeneity between cities  
suggests local context likely matter more than victim sex. Overall, the plot  
provides little consistent evidence that homicides with male victim are resolved  
at markedly different rates than those with female victim, however, in some  
cities (such as New York and Chicago) male-victim cases are actually less likely  
to be solved.

## Problem 2
```{r}
# Filter data
data <- weather_df %>%
        drop_na(tmax, tmin, prcp)
nrow(data)

# Fit model
fit_full <- lm(tmax ~ tmin + prcp, data = data)
full_r2 <- glance(fit_full)$r.squared
full_tmin <- tidy(fit_full) %>% 
             filter(term == "tmin") %>% 
             pull(estimate)
full_prcp <- tidy(fit_full) %>% 
             filter(term == "prcp") %>% 
             pull(estimate)
est_full <- tibble(param = c("r2", "beta_tmin", "beta_prcp"),
                   est = c(full_r2, full_tmin, full_prcp))

# Helper function
boot_once <- function(df) {
             idx <- sample.int(nrow(df), nrow(df), replace = TRUE)
             dfb <- df[idx, , drop = FALSE]
             fit <- lm(tmax ~ tmin + prcp, data = dfb)
             tibble(r2 = glance(fit)$r.squared,
                    beta_tmin = tidy(fit) %>% 
                    filter(term == "tmin") %>% 
                    pull(estimate), beta_prcp  = tidy(fit) %>% 
                    filter(term == "prcp") %>% 
                    pull(estimate))
}

loop <- 5000
result <- vector("list", loop)
for (i in seq_len(loop)) {
  result[[i]] <- boot_once(data)
}

# Bootstrap
boot <- bind_rows(result)
dim(boot)
head(boot)

# Confidence interval
ci_r2 <- quantile(boot$r2, c(0.025, 0.975), na.rm = TRUE)
ci_tmin <- quantile(boot$beta_tmin, c(0.025, 0.975), na.rm = TRUE)
ci_prcp <- quantile(boot$beta_prcp, c(0.025, 0.975), na.rm = TRUE)
ci_tbl <- tibble(param = c("r2", "beta_tmin", "beta_prcp"),
                 q2.5 = c(ci_r2[1], ci_tmin[1], ci_prcp[1]),
                 q97.5 = c(ci_r2[2], ci_tmin[2], ci_prcp[2]))
ci_tbl

boot_long <- boot %>%
             pivot_longer(everything(), names_to = "param", 
                          values_to = "value") %>%
             mutate(param = recode(param, r2 = "R-squared", 
                                   beta_tmin = "β (tmin)", 
                                   beta_prcp = "β (prcp)"))
# Estimates
est_full_plot <- est_full %>%
                 mutate(param = recode(param,r2 = "R-squared", 
                                       beta_tmin = "β (tmin)", 
                                       beta_prcp = "β (prcp)"))

# Plot of distribution
ggplot(boot_long, aes(x = value)) + geom_histogram(bins = 40, color = "white") +
geom_vline(data = est_full_plot, aes(xintercept = est), color = "red") + 
facet_wrap(~ param, scales = "free", ncol = 3) + 
labs(title = "Bootstrap distributions (5000 in total)", x = NULL, y = "Count")
```

Histograms of the 5000 estimates are unimodal and roughly symmetric. R^2 is  
tightly concentrated around about 0.94, indicating good fit. β_1 is centered a  
little above 1 (about 1.01–1.02), while β_2 is centered near −0.006, showing  
narrow spreads.

95% bootstrap percentile CIs (2.5%–97.5%) are R^2: [0.934, 0.946], β_1 (tmin):  
[1.006, 1.028] and β_2 (prcp): [−0.008, 0.004].

## Problem 3
```{r}
# Read data
bw_raw <- readr::read_csv("data/birthweight.csv", show_col_types = FALSE) 

# Process data
bw <- bw_raw %>%
      janitor::clean_names() %>%
      mutate(babysex = factor(babysex, levels = c(1, 2), 
                              labels = c("Male", "Female")),
             malform = factor(malform, levels = c(0, 1), 
                              labels = c("Absent", "Present")),
             mrace = factor(mrace, levels = c(1, 2, 3, 4, 8),
                            labels = c("White","Black","Asian",
                                       "Puerto Rican","Other")),
             frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9), 
                            labels = c("White","Black","Asian","Puerto Rican",
                                       "Other","Unknown")))
vars_use <- c("bwt", "bhead", "blength", "gaweeks", "babysex", "momage", 
              "ppbmi", "wtgain", "smoken", "parity")
bw_mod <- bw %>% 
          drop_na(all_of(vars_use))

# Fit model
mod_spec <- bwt ~ bhead + blength + gaweeks + babysex + momage + ppbmi + 
                  wtgain + smoken + parity
mod_main <- lm(mod_spec, data = bw_mod)

# Result
glance(mod_main) %>% 
select(r.squared, adj.r.squared, sigma)
tidy(mod_main) # coefficients and significance

# Model residuals
bw_result <- bw_mod %>%
             add_predictions(mod_main) %>%
             add_residuals(mod_main)

# Plot of model residuals against fitted values 
bw_result %>%
ggplot(aes(x = pred, y = resid)) + geom_point(alpha = 0.25) + 
geom_smooth(se = FALSE, method = "loess") + 
labs(title = "Residuals vs Fitted", x = "Fitted birthweight (g)", 
     y = "Residual (g)") + theme_minimal(base_size = 12)

# Cross-validation
cv_df <- crossv_mc(bw_mod, n = 100)  # 100 次随机切分，课程常用规模
cv_comp <- cv_df %>%
           mutate(m_main = map(train, ~ lm(mod_spec, data = as.data.frame(.x))),
                  m_a = map(train, ~ lm(bwt ~ blength + gaweeks, 
                                        data = as.data.frame(.x))),
                  m_b = map(train, ~ lm(bwt ~ bhead * blength * babysex, 
                                        data = as.data.frame(.x))),
                  rmse_main = map2_dbl(m_main, test, ~ rmse(.x, 
                                                            as.data.frame(.y))),
                  rmse_a = map2_dbl(m_a, test, ~ rmse(.x, as.data.frame(.y))),
                  rmse_b = map2_dbl(m_b, test, ~ rmse(.x, 
                                                      as.data.frame(.y)))) %>%
           select(starts_with("rmse_")) %>%
           pivot_longer(everything(), names_to = "model", 
                        values_to = "rmse") %>%
           mutate(model = recode(model, rmse_main = "Proposed (multi-factor)",
                                 rmse_a = "Length + GA (main effects)",
                                 rmse_b = "Head*Length*Sex (all interactions)"))

# Summary
cv_summary <- cv_comp %>%
              group_by(model) %>%
              summarise(mean_rmse = mean(rmse), sd_rmse = sd(rmse), 
                        .groups = "drop") %>%
              arrange(mean_rmse)
cv_summary

# Plot of 3 models' rmse
cv_comp %>%
ggplot(aes(x = model, y = rmse)) + geom_violin(fill = "grey95") + 
geom_boxplot(width = 0.15, outlier.alpha = 0.2) + coord_flip() +
labs(title = "Cross-validated RMSE by model (n = 100 splits)", x = NULL, 
     y = "RMSE (grams)") + theme_minimal(base_size = 12)
```

The following linear model is proposed (where bwt is in grams): bwt ∼ bhead +  
blength + gaweeks + babysex + momage + ppbmi + wtgain + smoken + parity. bhead,  
blength and gaweeks directly reflect the fetal body size and gestational age,  
and are the core determinants of the fetal weight; babysex reflects gender  
differences; momage, ppbmi, wtgain reflect the maternal nutritional status;   
smoken reflects the effect of smoking during pregnancy and parity reflects the  
possible influence of previous childbirth history on body weight.  
The result of this model on the data is R^2 = 0.699 (after adjustment 0.699),  
and the residual standard deviation is approximately 281 g, indicating that  
the model can explain about 70% of the weight variation. Within the main fitting  
range (approximately 2600–3800 g), the residuals are symmetrically distributed  
around 0 and no systematic bias is observed. 

Comparing the prediction errors (RMSE in grams) of the three models, the  
proposed multi-factor model is mean 282 g (SD 9.7), minimum; the model of Head *  
Length * Sex with all interactions included is mean 289 g (SD 10.7), slightly  
lower; and the model of only the main effect of Length + GA is mean 333 g  
(SD 14.8), the worst of all.